---
title: "covid 19"
author: "Dr K"
date: "5/2/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```

```{r eval=FALSE}
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
```

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ There is no correlation between sex and dying.

$H_A$ There is a correlation between sex and dying.


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```



## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


The ggplot barchart shows that there were more females than males observed. There were more males than died then females. About 65% those that died were male and there were less males observed in the study.

  
## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The results from the bottom table shows 1.4% of females died, while 3% of males died. Double the amount of males died than females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test tells us that the men were 2.3 times as likely to die than the females in this study. The p-value reveals that the chances of getting this result if there was no difference is 0.0009.

## Conclusions

The results from the barchart showed 65% of those that died being male with only 35% being female. The numerical results showed that 3% of males died and 1.4% of females died. The fisher and p-value text told us men were 2.3 times as likely to die with the chances of seeing these results with no difference in 0.0009. These results show that death is not independent of sex.
